EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-21 03_edipack2.tex
Version 1.0 completed. Ready to share with all authors. Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

{\bf REMARK TODO}
Notes on the norm function: a) the definition for conventional
matrices, ;b) the definition for the replica/general case where you
can choose between different norm defition.
\fixme{There's no actual reason to not provide the ``elemental" norm for
the normal and hybrid bath parametrizations, as it is relevant or not
depending on the off-diagonal structure of $\Theta(z)$, which depends
only on \texttt{ed\_mode}, not \texttt{bath\_type}. What do we do? I would
vote for describing what the different norms do and vaguely mention that
they do not work for all possible \texttt{bath\_type} values, without 
entering any detail on the reason why. Or even not mention it, and slowly
start to update the code to make it work with all possible combinations...}
%GBtodo placeholder

>>>>

{\bf REMARK TODO}
Notes on the norm function: a) the definition for conventional
matrices, ;b) the definition for the replica/general case where you
can choose between different norm defition.
\fixme{After careful inspection, I (GB) realize that the elemental norm 
was never really ported from CDFMT-ED to EDIpack (there's no different
matrix weight than all ones, which is the standard chi2, of course). 
Hence the \texttt{cg\_norm} options is really not doing anything significant,
as with "elemental" it gives the usual $\chi^2$ (or $\chi^q$, if you change the
\texttt{cg\_pow}...) and for "Frobenius" it gives again the usual $\chi^2$,
unless you put $q\neq2$, which gives an ill-defined matrix norm...
(it mixes an effective $q=2$ coming from the matrix multiplication with 
$q\neq2$ when you do the explicit $q$-root). I think we
should just get rid of \texttt{cg\_norm} and do the traditional $\chi^q$, as 
indeed we are already doing by default. Implemented like this, it is just random
no-consequence code.}
%GBtodo placeholder

### 2025-05-21 05_examples.tex
Version 1.0 completed. Ready to share with all authors. Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\begin{figure}%[t!]
  \includegraphics[width=\linewidth]{figures/figBetheW2D.pdf}
    \caption{\label{figEx1W}%
        }
\end{figure}

\subsubsection{Finite temperature solution (W2Dynamics interface)}\label{SecExamplesBetheDMFTw2d}
To demonstrate use of the w2dynamics interface, here we briefly
discuss how to solve the DMFT problem for the Bethe lattice using
W2Dynamics interface. \fixme{For definitiness we compare the finite low
temperature results obtained via the two solvers: ED and CTQMC.  }

Unlike when the Fortran API is used, none of the
five steps of the iterative algorithm listed there are left to the
user to implement themselves. Instead, in addition to possibly
non-default values for \NAME input variables, the model specification
must also be added to a w2dynamics parameter file and depending on the
model, further input files might need to be provided. Further
differences are that initialization is performed in w2dynamics by
starting from a self-energy of zero or reading a self-energy value
from an old file, which results in the initial Weiss field that the
initial bath parameters are fitted to, and that automatic convergence
checking is not currently available. Finally, using the same settings
as for the Fortran example we arrive at the w2dynamics parameter file
shown in the following listing.

\begin{lstlisting}[style=fstyle,language={},numbers=none,basicstyle={\scriptsize\ttfamily}]
[General]
DOS             = Bethe           # support for the Bethe lattice is built-in
half-bandwidth  = 1               # list of half-bandwidths per orbital
NAt             = 1               # number of impurities
beta            = 1000            # inverse temperature, defines Matsubara grid spacing
mu              = 1.0             # chemical potential set to achieve half-filling
EPSN            = 0.0             # turns off filling-based chemical potential search
DMFTsteps       = 100             # given no convergence checking, we might want fewer
magnetism       = para            # symmetrize self-energies per spin
FileNamePrefix  = bethe_dmft_U2   # prefix for the output file name
fileold         = bethe_dmft*hdf5 # file to read an initial self-energy from
readold         = 0               # iteration number to read initial self-energy from, 0 turns off
mixing          = 0.5             # mixing, but mixes self-energies and not Weiss fields
mixing_strategy = linear          # linear mixing as in the Fortran example
solver          = EDIPACK         # use EDIpack2 as impurity solver, not default CTHYB

@@

[EDIPACK]                         # further non-default parameters, also taken from Fortran example
NBATH           = 7               # number of bath sites
ED_TWIN         = True            # use twin symmetry
LFIT            = 2048            # number of Matsubara frequencies used for the bath fit
LANC_NGFITER    = 500             # number of Lanczos iterations for Green's function
CG_FTOL         = 1e-10           # conjugate-gradient tolerance
CG_NITER        = 2048            # maximum number of conjugate-gradient iterations
\end{lstlisting}

With this parameter file, we can use the {\tt DMFT.py} script of
w2dynamics to perform the execution and find the results in form of
both a typical w2dynamics HDF5 output file whose name starts with {\tt
  bethe\textunderscore{}dmft\textunderscore{}U2} and a number of folders whose names start with the
name of the HDF5 output file that contain unavoidable or explicitly
requested \NAME input and output files for each individual DMFT
iteration. We can extract results from the HDF5 output file e.g. using
the {\tt hgrep} script of w2dynamics. The following listing shows how
to perform a calculation with 10 parallel MPI processes and print the
impurity self-energy result of the last DMFT iteration to the terminal
assuming that the parameter file is {\tt bethe\textunderscore{}dmft.in} in the working
directory and w2dynamics installed to {\tt /opt/w2dynamics}.

\begin{lstlisting}[style=mybash]
mpiexec -n 10 /opt/w2dynamics/DMFT.py bethe_dmft.in
python /opt/w2dynamics/hgrep latest siw-full -1
\end{lstlisting}

>>>>

\subsubsection{Finite temperature solution (W2Dynamics interface)}\label{SecExamplesBetheDMFTw2d}
To demonstrate the capabilities of the W2Dynamics interface to \NAME,  here we briefly discuss how to solve the same problem, i.e. the Hubbard model on the Bethe lattice within DMFT W2Dynamics. 
In order to showcase the capabilities of the \NAME to address low-temperature problems comaring the W2Dynamics  Continuous-Time Quantum Monte Carlo (CTQMC) with hybridization expansion method against the \NAME solver at finite temperature.

As we have seen the \NAME API, enabling access to the impurity solver and bath optimization procedures, requires the user to implement (using different methods) the DMFT algorithm.  W2Dynamics offers a structurally different solution, providing a single Python script {\tt DMFT.py} which takes care of the whole DMFT calculation via dedicated classes to implement generic self-consistency. The W2Dynamics is then entirely controlled by model-dependent parameters file {\tt Parameters.in} which contains a number of variable specifications, including non-default ones controlling the ED solver inherited from \NAME interface. Further information about the functioning of W2Dynamics can be found in Ref.\onlinecite{W2DynWallerberger}           
The two methods have another important difference. 
The first concerns the initial point of the iterative DMFT solution algorithm. 
While \NAME starts from a given discrete bath, W2Dynamics is initialized with a zero Self-energy function or, in alternative, read this quantity from a converged solution. When using \NAME ED interface, the initial Weiss is determined using self-consistency and finally an initial discrete bath is obtained through the \NAME bath optimization procedure.

The W2Dynamics configuration file for the solution of the Bethe lattice problem at finite temperature  reads: 
\begin{lstlisting}[style=mybash,language={},numbers=none,basicstyle={\scriptsize\ttfamily}]
[General]
DOS             = Bethe           # support for the Bethe lattice is built-in
half-bandwidth  = 1               # list of half-bandwidths per orbital
NAt             = 1               # number of impurities
beta            = 1000            # inverse temperature, defines Matsubara grid spacing
mu              = 1.0             # chemical potential set to achieve half-filling
EPSN            = 0.0             # turns off filling-based chemical potential search
DMFTsteps       = 100             # given no convergence checking, we might want fewer
magnetism       = para            # symmetrize self-energies per spin
FileNamePrefix  = bethe_dmft_U2   # prefix for the output file name
fileold         = bethe_dmft*hdf5 # file to read an initial self-energy from
readold         = 0               # iteration number to read initial self-energy from, 0 turns off
mixing          = 0.5             # mixing, but mixes self-energies and not Weiss fields
mixing_strategy = linear          # linear mixing as in the Fortran example
solver          = EDIPACK         # use EDIpack2 as impurity solver, not default CTHYB

@@

[EDIPACK]                         # further non-default parameters, also taken from Fortran example
NBATH           = 7               # number of bath sites
ED_TWIN         = True            # use twin symmetry
LFIT            = 2048            # number of Matsubara frequencies used for the bath fit
LANC_NGFITER    = 500             # number of Lanczos iterations for Green's function
CG_FTOL         = 1e-10           # conjugate-gradient tolerance
CG_NITER        = 2048            # maximum number of conjugate-gradient iterations
\end{lstlisting}
Following W2Dynamics  specifics, we run the {\tt DMFT.py} script as:
\begin{lstlisting}[style=mybash,numbers=noneccd ]
mpiexec -n 10 /opt/w2dynamics/DMFT.py bethe_dmft.in
python /opt/w2dynamics/hgrep latest siw-full -1
\end{lstlisting}
Where, as for W2Dyanamics, the results are collected into a HDF5 archive, including the output files inherited from \NAME which is inspected using the provided script {\tt hgrep}. In this example we extract the Matsubara self-energy function $\Sigma(i\omega_n)$ for the last DMFT iteration.

\begin{figure}%[t!]
  \includegraphics[width=\linewidth]{figures/figBetheW2D.pdf}
    \caption{\label{figEx1W}%
    \textbf{Finite temperature DMFT solution.}
    Comparison of the imaginary parts of the Green's function $\Im{G}(i\omega_n)$ and Self-energy $\Im{\Sigma}(i\omega_n)$ from different solutions of the Hubbard model on the Bethe lattice using DMFT. Data are for $T/D=0.01$ and $U/D=2.0$. The results are obtained using original W2Dynamics CT-QMC solver and compared against the \NAME solver, either independently or using the W2Dynamics interface. 
        }
\end{figure}

In \figu{figEx1W} we report a comparison of the results obtained using the hybridization expansion CTQMC solver in W2Dynamics, the \NAME ED method available as solver W2Dynamics as well as independent calculation (Fortran API). Using DMFT, we solve the Hubbard model on a Bethe lattice for $T/D=0.01$ and $U=2.00$, corresponding to a correlated metallic state. 
In panel (A) we show the behavior of the imaginary part of the impurity Green's function $\Im{G}(i\omega_n)$ (note that W2Dynamics does not have access to real-axis Green's function). In panel (B) we show the same comparison for $\Im{\Sigma}(i\omega_n)$. 
The results obtained with the three methods are nearly identical already for a small number of bath levels ({\tt NBATH=7}) or low QMC statistics.
